# Predictions in generalized linear models

ideas taken from (this blog post)[https://www.fromthebottomoftheheap.net/2018/12/10/confidence-intervals-for-glms/]

## packages

```{r}
library(tidyverse) # for ggplot
```

## Some made-up data

```{r}
d=tribble(
  ~dose, ~status,
  1, "lived",
  2, "died",
  3, "lived",
  4, "lived",
  5, "died",
  6, "lived",
  7, "died",
  8, "died",
  9, "died"
)
d
```

```{r}
factor(d$status)
```

```{r}
fct_inorder(d$status)
```


## fit logistic regression

```{r}
d.1=glm(factor(status)~dose, family="binomial", data=d)
summary(d.1)
```

probability of *living* goes down as dose goes up

## predictions (for the observed doses)

```{r}
p=predict(d.1)
cbind(d, p)
```

these are predicted log-odds of living

Can make it into predicted probabilities:

```{r}
p=predict(d.1, type="response")
p
cbind(d,p)
```

Can also get standard errors:

```{r}
p=predict(d.1, type="response", se.fit = T)
p
cbind(d,p) -> preds
preds
```

Obvious thing, plus/minus twice the SE, makes no sense here:

```{r}
preds %>% mutate(lo=fit-2*se.fit, hi=fit+2*se.fit) -> preds2
preds2
```

plot predicted probs

```{r}
g <- ggplot(preds, aes(x = dose, y = fit)) +
    geom_line()
g
```

add wrong ci limits as ribbon

```{r}
g + geom_ribbon(data = preds2, aes(ymin = lo, ymax = hi),
                  alpha = 0.1) +
  geom_hline(yintercept=1, linetype="dashed")+
  geom_hline(yintercept=0, linetype="dashed")

```

## Doing it better

Work with *linear predictor*, and then invert it

```{r}
p=predict(d.1, se.fit=T)
cbind(d, p) %>% 
  mutate(lo=fit-2*se.fit, hi=fit+2*se.fit) -> preds3
preds3
```

getting the inverse link function

```{r}
fam=family(d.1)
str(fam)
```

there is a function `linkinv` that we can grab:

```{r}
inv_link=fam$linkinv
inv_link(0)
inv_link(1)
inv_link(-1)
```

work out inverse logit of predictions and hi and lo

```{r}
preds3
```

```{r}
preds3 %>% mutate(fit=inv_link(fit),
                  lo=inv_link(lo),
                  hi=inv_link(hi)) -> preds4
preds4
```

```{r}
g + geom_ribbon(data = preds4, aes(ymin = lo, ymax = hi),
                  alpha = 0.1) +
  geom_hline(yintercept=1, linetype="dashed")+
  geom_hline(yintercept=0, linetype="dashed")

```

